U.S. Labour Market Resilience

### 2024-03-18 paper/paper.qmd
Update new dataset

---
title: "U.S. Labour Market Resilience"
subtitle: "Analysis of Economic Crises on Industry Work Hours in 1998, 2008 and 2020 US General Social Surveys"
author: 
  - Mingjia Chen
  - Catherine Punnoose
  - Tianen (Evan) Hao
thanks: "Code and data are available at: https://github.com/ShadyEvan4830/Labour_Market_Resilience.git"
date: March 11, 2024
date-format: long
abstract: "Economic crises have impacted labour market working patterns and employment levels, but are all groups within the labour market suffering from similar trends? Focusing on the Asian Financial Crisis, the Global Financial Crisis, and the COVID-19 pandemic, this study uses data from the U.S. General Social Survey (GSS) to investigate the impact of major economic downturns on U.S. labour force dynamics, focusing specifically on “hours worked last week.” Our analysis reveals that economic downturns lead to a decline in average working hours, and identifies disparities in the effects of economic crises on demographic groups, with women and younger workers disproportionately affected. These findings underscore the necessity for targeted interventions to support the most vulnerable groups in the labour market."
format: pdf
toc: true
number-sections: true
bibliography: bibliography.bib
---

@@

Initial findings show disparities: younger workers, those with less education, and women face significant work-hour reductions during crises. This underscores the need for targeted policy interventions to address these disparities and strengthen labour market resilience.

@@

```{r}
#| label: fig-cate-comparisons
#| fig-cap: Categorical Working Hour Comparison between the Years
#| fig-subcap: ["Asian Financial Crisis (1998)","World Financial Crisis (2008)","During Covid-19 Pandemic (2021)", "Total"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

@@

Data_1998 <- select(data,work_hours,`1998`)
ggplot(Data_1998,aes(x=work_hours,y=`1998`)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Hours of Working", y = "Number of Reports") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

Data_2008 <- select(data,work_hours,`2008`)
ggplot(Data_2008,aes(x=work_hours,y=`2008`)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Hours of Working", y = "Number of Reports") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

Data_2021 <- select(data,work_hours,`2021`)
ggplot(Data_2021,aes(x=work_hours,y=`2021`)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Hours of Working", y = "Number of Reports") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

Data_Total <- select(data,work_hours,Total)
ggplot(Data_Total,aes(x=work_hours,y= Total)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Hours of Working", y = "Number of Reports") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

@@

## Average Working Hour Comparison Between Financial Events

@@

```{r}
#| label: fig-ave-hour-comparison
#| fig-cap: Average Working Hours Comparison between Financial Events
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-hour-working-percentage
#| fig-cap: Average Working Hours Percentages 
#| echo: false
#| warning: false
#| message: false
#| 
data <- read.csv(here::here("data/analysis_data/cleaned_categorized_data.csv"))
data <- data[c(2:ncol(data))]
# Get rid of the first letter X for each of the years
colnames(data) <- gsub("X", "", colnames(data))
work_hours <- rep(data$work_hours,35)
years <- colnames(data)[2:36]
years <- rep(years,each=7)

@@

@@

Non-response rates we illustrated in @fig-nonresponse-rate, in addition, could help indicate a significant difference in participants and data collection between the years. The year 1972, for example, is evidence of the gap in the data set; even though it is included in the data set, the non-response rate is 100%, meaning that no answers to this question were collected during that year. The non-response rate in general is similar among the years, not indicating significant differences between the years even during the financial crisis.

>>>>

---
title: "U.S. Labour Market Resilience"
subtitle: "Analysis of Economic Crises on Industry Work Hours in 1998, 2008 and 2020 US General Social Surveys"
author: 
  - Mingjia Chen
  - Catherine Punnoose
  - Tianen (Evan) Hao
thanks: "Code and data are available at: https://github.com/ShadyEvan4830/Labour_Market_Resilience.git. Thank Professor Alexander so much for the fantastic feedback."
date: March 11, 2024
date-format: long
abstract: "Economic crises have impacted labour market working patterns and employment levels, but are all groups within the labour market suffering from similar trends? Focusing on the Asian Financial Crisis, the Global Financial Crisis, and the COVID-19 pandemic, this study uses data from the U.S. General Social Survey (GSS) to investigate the impact of major economic downturns on U.S. labour force dynamics, focusing specifically on “hours worked last week.” Our analysis reveals that economic downturns lead to a decline in average working hours, and identifies disparities in the effects of economic crises on demographic groups, with women and younger workers disproportionately affected. These findings underscore the necessity for targeted interventions to support the most vulnerable groups in the labour market."
format: pdf
toc: true
number-sections: true
bibliography: bibliography.bib
---

@@

Initial findings show disparities: workers faced significant reductions in working hours during the crisis, with studies by [@Chen_Yeh_2021] and [@Singh_Shirazi_Turetken_2022] showing that women face a more severe job impact due to their unique social obligations. This highlights the need for targeted policy interventions to address these differences and enhance labour market resilience.

@@

```{r}
#| label: fig-cate-comparisons
#| fig-cap: Categorical Working Hours of last Week Comparison between the Years with Critial Financial Crisis
#| fig-subcap: ["Asian Financial Crisis (1998)","World Financial Crisis (2008)","During Covid-19 Pandemic (2021)", "Total"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

@@

Data_1998 <- select(data,work_hours,`1998`)
ggplot(Data_1998,aes(x=work_hours,y=`1998`)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Hours of Working Last Week", y = "Number of Responses") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

Data_2008 <- select(data,work_hours,`2008`)
ggplot(Data_2008,aes(x=work_hours,y=`2008`)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Hours of Working Last Week", y = "Number of Responses") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

Data_2021 <- select(data,work_hours,`2021`)
ggplot(Data_2021,aes(x=work_hours,y=`2021`)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Hours of Working Last Week", y = "Number of Responses") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

Data_Total <- select(data,work_hours,Total)
ggplot(Data_Total,aes(x=work_hours,y= Total)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Hours of Working Last Week", y = "Number of Responses") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

@@

## Average Working Hour Comparison Between Financial Crisis

@@

```{r}
#| label: fig-ave-hour-comparison
#| fig-cap: Average Working Hours Comparison between Financial Crisis
#| echo: false
#| warning: false
#| message: false

@@

As we illustrated in @fig-nonresponse-rate, non-response rates may also help indicate significant differences in participants and data collection over the years. For example, although we can still see a slight increase from 1992 to 2022 (no data for 1992), this suggests that a gradual rise in non-response rates from 1992 to 2022 may signal a decline in participant engagement and accessibility Variety. We speculate that this trend may be attributed to the rise of digital communications, privacy concerns, and survey fatigue. Demographic changes and technological advances, such as the widespread use of mobile devices, also play a role. Therefore, adapting survey methods to these changes is critical to maintaining data quality and representativeness.

```{r}
#| label: fig-hour-working-percentage
#| fig-cap: Average Working Hours Percentages 
#| echo: false
#| warning: false
#| message: false

data <- read.csv(here::here("data/analysis_data/cleaned_categorized_data.csv"))
data <- data[c(2:ncol(data))]
# Get rid of the first letter X for each of the years
colnames(data) <- gsub("X", "", colnames(data))
work_hours <- rep(data$work_hours,35)
years <- colnames(data)[2:36]
years <- rep(years,each=7)

@@

# Reference Helper Function code from 
# https://github.com/christina-wei/Financial-Wellness-in-US/blob/main/scripts/04-helper_functions.R

@@

```{r}
#| label: fig-demographic-comparison
#| fig-cap: Working Hours Comparisons with Demographic factors
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

sup_data <- read.csv(here::here("data/analysis_data/supplementary_data.csv"))
